The large scale clustering in dispersion measure space

### 2015-05-14 writeup/dm_space_distortions.tex
Some writeup of systematics.

@@

    The recently discovered fast radio bursts are a new class of radio
    transient that, due to their high dispersion from intervening plasma,
    have been interpreted as
    originating from cosmological distances.  The rate of fast radio bursts
    has been estimated to be as high as 10\,000 per sky per day.  If
    cosmological, these events should trace the large-scale structure of the
    Universe in some way. Furthermore, because the dispersion measure (DM) can be
    used as a proxy for radial distance (like redshift),
    clustering can be studied in three dimensions. Due to inhomogeneities in
    the plasma's density dispersion measure is an imperfect proxy or radial
    distance and we show that this leads to additional clustering terms---the
    dispersion measure space distortions. We calculate the DM space
    power spectra for a toy model of electron and FRB clustering and show that
    the clustering signal could be detected in a survey of \red{10000} FRB
    events, as is expected to be produced by the Canadian Hydrogen Intensity
    Mapping Experiment.

@@

\section{Introduction}

\section{Density perturbations in dispersion measure space}

The dispersion measure of a signal observed in some angular direction $\hat{n}$ and
originating from co moving radial distance $\chi$ is
\begin{equation}
    {\rm DM}(\hat{n}, \chi) = \int_0^{\chi} \ud\chi' a(\chi)^2
        n_e(\hat{n}\chi', \chi').
\end{equation}
Here
$n_e(\vec{x}, \chi)$ is the free electron density as a function of location and
conformal time. Note that we use $\chi$ as our radial distance and time
coordinate, as opposed to redshift which is more appropriate for spectroscopic
applications. We model the cosmological 
electron density as containing a homogeneous part
and perturbations
$n_e(\vec x, \chi) = \bar{n}_e(\chi) \left[1 + \delta_e(\vec x, \chi)\right]$.  \red{We take
all of the electrons to be in a diffuse ionized state such that
$\bar{n}_e(\chi) = \bar{n}_{e0} / a(\chi)^3$ (not necessary).} The dispersion
measure is thus
\begin{equation}
    {\rm DM}(\hat n, \chi) = \int_0^\chi \ud\chi' a(\chi')^2 \bar{n}_e(\chi')
       \left[1 + \delta_e(\hat{n}\chi',\chi')\right].
\end{equation}

@@

Combining the above equations (keeping only terms first order in the density
perturbations), we find that
\begin{equation}
\frac{\ud \chi_s}{\ud \chi} = (1 + \delta_e(\hat n \chi, \chi),
\end{equation}
and thus
\begin{equation}
\chi_s - \chi = \int_0^\chi \ud \chi' \delta_e(\hat n \chi').
\end{equation}

@@

\section{Two-point statistics in DM space}

Unlike in redshift-space distortions, Equation~\ref{e:delta_s} does not have a
simple form in harmonic space. The equation's third term couples harmonic modes
and thus the two-point statistics cannot be phrased as a simple power spectrum.
We will phrase the two point statistic in terms of $C^{ss}_\ell(\chi,\chi')$, which is
the cross-correlation angular power-spectrum of the DM space over-density
on shells at $\chi$ and $\chi'$.
\begin{align}
    &\delta_{\ell\ell'}\delta_{mm'}C^{ss}_\ell(\chi, \chi') = \nonumber \\
    & \quad
        \left\langle
        \int\ud\Omega Y_{\ell m}(\hat n)
       \delta_{fs}(\hat n \chi, \chi)
        \int\ud\Omega' Y_{\ell' m'}(\hat n')
        \delta_{fs}(\hat n' \chi', \chi')
        \right\rangle
\end{align}
The first two term in Equation~\ref{e:delta_s} are stationary. For stationary,
isotropic, tracers $x$ and $y$ we have 
$\langle \delta_x(\vec k, \chi) \delta_y(\vec k', \chi) \rangle = (2\pi)^3
\delta^3(\vec k - \vec k') P_{xy}(k, \chi)$.  If for a moment we ignore time
dependance, the angular cross power spectrum of such tracers is
\begin{align}
    %\delta_{\ell\ell'}\delta_{mm'}C^{xy}_\ell(\chi, \chi')
    %&= \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
    %    \langle \delta_x(\hat n \chi) \delta_y(\hat n' \chi') \rangle,
    %    \\
    C^{xy}_\ell(\chi,\chi')
    &= \frac{2}{\pi}
\int_0^\infty\ud k k^2 j_\ell(k\chi) j_{\ell}(k\chi')P_{xy}(k).
\end{align}
In reality the power spectrum evolves on the order of a Hubble time.
The angular cross-correlations will be very small unless $\chi$ and $\chi'$ are
within a few correlation lengths of one another, roughly a few hundred
megaparsecs.  The evolution of the power-spectrum is negligible over these time
differences which leads to a straight forward way to include the evolution of
the perturbations:
\begin{align}
C^{xy}_\ell(\chi,\chi') 
    \approx \frac{2}{\pi}
    \int_0^\infty\ud k k^2
    j_\ell(k\chi) j_{\ell}(k\chi')
    P_{xy}(k,(\chi + \chi')/2),\nonumber &\\
    |\chi - \chi'| \ll 1/aH.&
\end{align}

@@

\subsection{Modelling the angular power-spectrum}

@@

\section{Measurement}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%            Scrap           %%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{widetext}

\newpage

\section{Details not to be included}

We use the following conventions:
\begin{equation}
    P_{xy}(k, \chi) = b_x b_y g(\chi)^2 a(\chi)^2 P(k, 0)
\end{equation}

\begin{equation}
    \delta_x(\vec x, \chi) =  b_x g(\chi) a(\chi) \int\frac{\ud^3\vec k}{(2 \pi)^3}
        e^{i\vec k \cdot \vec x} \delta(\vec k, 0)
\end{equation}

\begin{equation}
    w^{xy}(\hat n \cdot \hat n', \chi, \chi')
    \equiv \langle \delta_x(\vec x) \delta_y(\vec x') \rangle
\end{equation}

For any statistically stationary tracers $x$ and $y$,
ignoring time dependance (redshift evolution)
of the power spectrum:
\begin{align}
    \delta_{\ell\ell'}\delta_{mm'}C^{xy}_\ell(\chi, \chi')
    &= \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
        \langle \delta^x(\hat n \chi) \delta^y(\hat n' \chi') \rangle
        \\
    &=
        \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
        \int\frac{\ud^3\vec k}{(2 \pi)^3} \frac{\ud^3\vec k'}{(2 \pi)^3}
        e^{i\vec k \cdot \vec x} e^{i\vec k' \cdot \vec x'}
        \langle \delta(\vec k) \delta(\vec k') \rangle
        \\
    &=
        \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
        \int\frac{\ud^3\vec k}{(2 \pi)^3} 
        e^{i\vec k \cdot (\vec x - \vec x')} P_{xy}(k)
        \\
    &=
        \int\frac{\ud^3\vec k}{(2 \pi)^3} P_{xy}(k)
        \int\ud\Omega Y_{\ell m}(\hat n)
        e^{i\vec k \cdot \vec x}
        \int\ud\Omega' Y_{\ell' m'}(\hat n')
        e^{-i\vec k \cdot \vec x'}
        \\
    &=
        \int\frac{\ud^3\vec k}{(2 \pi)^3} P_{xy}(k)
        \left[(4 \pi) i^\ell j_\ell(k\chi) Y_{\ell m}^*(\hat k)\right]
        \left[(4 \pi) i^{-\ell'} j_{\ell'}(k\chi') Y_{\ell' m'}(\hat
        k)\right]
        \\
    &=
        \frac{2}{\pi}
        \int_0^\infty\ud k k^2 j_\ell(k\chi) j_{\ell'}(k\chi')P_{xy}(k)
        \int\ud \Omega_k i^{(\ell - \ell')}
        Y_{\ell' m'}^*(\hat k) Y_{\ell m}(\hat k)
        \\
    &= \delta_{\ell\ell'}\delta_{mm'}\frac{2}{\pi}
        \int_0^\infty\ud k k^2  j_\ell(k\chi) j_{\ell}(k\chi')P_{xy}(k)
        \\
    C^{xy}_\ell(\chi,\chi') 
    &= \frac{2}{\pi}
\int_0^\infty\ud k k^2 j_\ell(k\chi) j_{\ell}(k\chi')P_{xy}(k)
\end{align}

It is convienient to reparameterize this in terms of
$\bar\chi \equiv (\chi + \chi') /2$ and $\Delta \chi \equiv \chi' - \chi$.
We expect the power spectrum to vary slowly as a function of $\bar\chi$ and
very quickly as a function of $\Delta\chi$. However it will be highly
suppressed for $\ell \gg \bar\chi/\Delta\chi$. These facts will be very
helpfull in subsequent numerical integrals.

Also since smallish value of $\Delta\chi$ are of interest, this leads us to
an obviouse way to include redshift dependance of the
power spectrum:
\begin{equation}
C^{xy}_\ell(\bar\chi,\Delta\chi) 
    = \frac{2}{\pi}
    \int_0^\infty\ud k k^2
    j_\ell(k\chi) j_{\ell}(k\chi')
    P_{xy}(k,\bar\chi) \qquad \Delta\chi \ll 1/aH
\end{equation}

Define $\delta_d$ as:
\begin{equation}
    \delta_d(\hat n \chi) = \int_0^\chi \ud \chi' \delta_e(\hat n \chi')
\end{equation}
which doesn't have the same units as over-density, but that is okay.
\begin{equation}
C^{dd}_\ell(\chi,\chi')
    =
    \frac{2}{\pi}
    \int_0^\chi\ud\chi''
    \int_0^{\chi'}\ud\chi'''
    \int_0^\infty\ud k k^2 j_\ell(k\chi'') j_{\ell}(k\chi'''
    )P_{ee}(k, (\chi''+\chi''')/2)
\end{equation}

For numerical integration, perform the following change of variables to make
things much easier.
\begin{align}
C^{dd}_\ell(\chi,\chi')
&=
    \frac{2}{\pi}
    \int_0^{(\chi' + \chi)/2}\ud\bar\chi'
    \int_{\max(-2\bar\chi', -2(\chi -\bar\chi'))}^{\min(2\bar\chi', 2(\chi' -
    \bar\chi'))}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\bar\chi' - \Delta'/2))
    j_{\ell}(k(\bar\chi' + \Delta'/2))
    P_{ee}(k, \bar\chi')
    \\
    &\approx
    \frac{4}{\pi}
    \int_0^{\chi_m}\ud\bar\chi'
    \int_{0}^{\infty}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\bar\chi' - \Delta'/2))
    j_{\ell}(k(\bar\chi' + \Delta'/2))
    P_{ee}(k, \bar\chi')
\end{align}
At this point we could again to the transformation to $C_{\ell\chi\chi'}$
leaving the expression as a triple integral. This is not as computationally
costly as might be expected at first glance, since the $\chi''$ and $\chi'''$
integrals are almost the same for all $\chi$, $\chi'$ pairs with only the
limits of integration changing. Thus the full set of $C_{\ell\chi\chi'}$'s can
be computed with a cumulative sum, only performing one new integral (over $k$)
for each element.

Finally, the cross terms:
\begin{equation}
C^{dx}_\ell(\chi,\chi')
    =
    \frac{2}{\pi}
    \int_0^\chi\ud\chi''
    \int_0^\infty\ud k k^2 
    j_\ell(k\chi') j_{\ell}(k\chi'')P_{ex}(k, (\chi' + \chi'')/2)
\end{equation}

with the following change of variables for numerical integration:
\begin{align}
C^{dx}_\ell(\chi,\chi')
    =&
    \frac{2}{\pi}
    \int_{-\chi'}^{\chi - \chi'}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k\chi')
    j_{\ell}(k(\chi'+\Delta'))
    P_{ex}(k, \chi' + \Delta'/2)
    \\
C^{dx}_\ell(\chi,\chi') + C^{xd}_\ell(\chi,\chi')
    \approx&
    \frac{2}{\pi}
    \int_{-\chi'}^{\chi - \chi'}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\chi'-\Delta'/2))
    j_{\ell}(k(\chi'+\Delta'/2))
    P_{ex}(k, \chi')
    \nonumber\\
    &+
    \frac{2}{\pi}
    \int_{-\chi}^{\chi' - \chi}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\chi-\Delta'/2))
    j_{\ell}(k(\chi+\Delta'/2))
    P_{ex}(k, \chi)\\
    \approx&
    \frac{\red{4}}{\pi}
    \int_{0}^{\infty}\ud\Delta'
    \int_0^\infty\ud k k^2 j_\ell(k(\chi_m-\Delta'/2))
    j_{\ell}(k(\chi_m+\Delta'/2))
    P_{ex}(k, \chi_m)
\end{align}
The second line exploits the fact that for $\ell \gg 1$, $\chi \gg \chi'/\ell$ and
$\chi' \gg \chi/\ell$, only $\Delta'\ll\chi$ contributes significantly to the
integral. Also the expressions are very weak functions of $\chi$ and thus
may be shifted by $\Delta/2$ without incurring much error. The third line
exploits the fact that only one term contributes if $\chi$ and $\chi'$ are
widely separated, and if they are close then the two terms combine to the given
expression.

\red{Actually, with the approximations we've made, we can actually do the
$\Delta$ integrals analytically by transferring all the $\Delta$ to a single
Bessel and shifting the power spectrum to be independent of $\Delta$.  Should
be pretty good for $\ell > 10$.  That is nice because it will control all our
numerical issues with oscillating Bessel functions.}

Assembling terms:
\begin{align}
C^{ss}_\ell(\chi,\chi') = &
    \frac{2}{\pi}
    \int_0^\infty\ud k k^2
    j_\ell(k\chi) j_{\ell}(k\chi')
    \left[ P_{ff}(k, \bar\chi)
    + P_{ee}(k, \bar\chi)
    - 2P_{fe}(k, \bar\chi) \right]
    \nonumber\\
    & +
    \frac{2}{\pi}
    \left(\frac{1}{\bar{n}_f}\frac{\ud \bar{n}_f}{\ud \chi}
    + \frac{2}{\chi} \right)
    \left(\frac{1}{\bar{n}_f}\frac{\ud \bar{n}_f}{\ud \chi'}
    + \frac{2}{\chi'} \right)
    \int_0^\chi\ud\chi''
    \int_0^{\chi'}\ud\chi'''
    \int_0^\infty\ud k k^2 j_\ell(k\chi'') j_{\ell}(k\chi''')
    P_{ee}(k, (\chi''+\chi''')/2)
    \nonumber\\
    & +
    \frac{2}{\pi}
    \left(\frac{1}{\bar{n}_f}\frac{\ud \bar{n}_f}{\ud \chi}
    + \frac{2}{\chi} \right)
    \int_0^\chi\ud\chi''
    \int_0^\infty\ud k k^2 
    j_\ell(k\chi') j_{\ell}(k\chi'')
    \left[ P_{ee}(k, (\chi' + \chi'')/2)
    - P_{fe}(k, (\chi' + \chi'')/2) \right]
    \nonumber\\
    & +
    \frac{2}{\pi}
    \left(\frac{1}{\bar{n}_f}\frac{\ud \bar{n}_f}{\ud \chi'}
    + \frac{2}{\chi'} \right)
    \int_0^{\chi'}\ud\chi''
    \int_0^\infty\ud k k^2 
    j_\ell(k\chi) j_{\ell}(k\chi'')
    \left[ P_{ee}(k, (\chi + \chi'')/2)
    - P_{fe}(k, (\chi + \chi'')/2) \right]
\end{align}

\section{Small angles and Limber}

Alternately, on the flat sky and for small angles:
\begin{align}
(2\pi)^2\delta^2(\vec \ell - \vec \ell')C_\ell(\chi, \chi')
    &=
        \int\ud^2\vec\theta\ud^2\vec\theta'
        e^{-i\vec\ell\cdot\vec\theta} e^{-i\vec\ell'\cdot\vec\theta'}
        \langle \delta(\hat n \chi) \delta(\hat n' \chi') \rangle
        \\
    &\approx
        \int\ud^2\vec\theta\ud^2\vec\theta'
        e^{-i\vec\ell\cdot\vec\theta} e^{-i\vec\ell'\cdot\vec\theta'}
        \int\frac{\ud^3\vec k}{(2 \pi)^3} 
        e^{i\vec k \cdot (\vec x - \vec x')} P(k, \bar\chi)
        \\
    &=
        \int\frac{\ud^3\vec k}{(2 \pi)^3} 
        e^{i k_\parallel (\chi - \chi')} P(k, \bar\chi)
        \int\ud^2\vec\theta e^{-i(\vec\ell-\bar\chi\vec k_\bot)\cdot\vec\theta}
        \int\ud^2\vec\theta' e^{-i(\vec\ell'-\bar\chi\vec k_\bot)\cdot\vec\theta'}
        \\
    &=
        \int\frac{\ud^3\vec k}{(2 \pi)^3} 
        e^{i k_\parallel (\chi - \chi')} P(\sqrt{k_\parallel^2 + k_\bot^2}, \bar\chi)
        (2\pi)^4 \delta^2(\vec\ell - \bar\chi \vec k_\bot)
        \delta^2(\vec\ell' - \bar\chi\vec k_\bot)
        \\
    &=
        \int_{-\infty}^\infty\frac{\ud k_\parallel}{(2 \pi)} 
        e^{i k_\parallel (\chi - \chi')}
        P(\sqrt{k_\parallel^2 + \ell^2/\bar\chi^2}, \bar\chi)
        \frac{(2\pi)^2\delta^2(\vec\ell' - \vec \ell)}{\bar\chi^2}
        \\
C_\ell(\chi, \chi')
    &= \frac{1}{\bar\chi^2}
        \int_{-\infty}^\infty\frac{\ud k_\parallel}{(2 \pi)} 
        e^{i k_\parallel (\chi - \chi')}
        P(\sqrt{k_\parallel^2 + \ell^2/\bar\chi^2}, \bar\chi)
\end{align}
Which is valid for $\ell \gg 1$ and $\chi - \chi' \ll \bar\chi$.
$\bar \chi \equiv (\chi + \chi') /2$

Or for flat sky and small angles:
\begin{align}
C^{dd}_\ell(\chi,\chi') 
    &=
    \int_0^\chi\ud\chi''
    \int_0^{\chi'}\ud\chi'''
    \frac{1}{\bar\chi^{\prime 2}}
    \int_{-\infty}^\infty\frac{\ud k_\parallel}{(2 \pi)} 
    e^{i k_\parallel (\chi'' - \chi''')}
    P_{ee}(\sqrt{k_\parallel^2 + \ell^2/\bar\chi^2}, (\chi'' + \chi''')/2)
    \\
    &\approx
    \int_0^\chi\ud\chi''
    \int_0^{\chi'}\ud\chi'''
    \frac{1}{\bar\chi^{\prime 2}}
    \int_{-\infty}^\infty\frac{\ud k_\parallel}{(2 \pi)} 
    e^{i k_\parallel (\chi'' - \chi''')}
    P_{ee}(\ell/\bar\chi', (\chi'' + \chi''')/2)
    \\
    &=
    \int_0^\chi\ud\chi''
    \int_0^{\chi'}\ud\chi'''
    \frac{1}{\bar\chi^{\prime 2}}
    \delta(\chi'' - \chi''')
    P_{ee}(\ell/\bar\chi', (\chi'' + \chi''')/2)
    \\
    &=
    \int_0^{\min(\chi,\chi')}\ud\chi''
    \left(\frac{1}{\chi''}\right)^2
    P_{ee}(\ell/\chi'', \chi'')
\end{align}
where the second line applies Limber's approximation. 
$\bar\chi' \equiv (\chi'' + \chi''')/2$.

Or under Limber:
\begin{align}
C^{dx}_\ell(\chi,\chi')
    &=
    \int_0^\chi\ud\chi''
    \frac{1}{\bar\chi^{\prime 2}}
    \delta(\chi'' - \chi')
    P_{ex}(\ell/{\bar\chi'}, (\chi'' + \chi')/2)
    \\
    &=
    \left(\frac{1}{\chi'}\right)^2
    P_{ex}(\ell/\chi', \chi')
    \qquad \textrm{if $\chi' < \chi$ and 0 otherwise.}
%    \\
%C^{dx}_\ell(\chi,\chi') + C^{xd}_\ell(\chi,\chi')
%    &=
%    \left(\frac{1}{\chi^m}\right)^2
%    P_{ex}(\ell/\chi^m, \chi_m)
\end{align}

Accumulating all terms for Limber and assuming a linear bias model:
\begin{align}
C^{ss}_\ell(\chi, \chi') =&~
    \left(\frac{b_f - b_e }{\bar\chi}\right)^2
        \int_{-\infty}^\infty\frac{\ud k_\parallel}{(2 \pi)} 
        e^{i k_\parallel \Delta\chi}
        P(\sqrt{k_\parallel^2 + \ell^2/\bar\chi^2}, \bar\chi)
    \nonumber \\ &+
    \left(\frac{1}{\bar{n}_f}\frac{\ud \bar{n}_f}{\ud \chi_m}
    + \frac{2}{\chi_m} \right)^2
    b_e^2 \int_0^{\chi_m}\ud\chi''
    \frac{1}{\chi^{\prime\prime 2}}
        P(\ell/\chi'',\chi'')
    \nonumber \\ &+
    \left(\frac{1}{\bar{n}_f}\frac{\ud \bar{n}_f}{\ud \chi_m}
    + \frac{2}{\chi_m} \right)
    \frac{b_e(-b_f + b_e)}{\chi_m^2}
        P(\ell/\chi_m, \chi_m)
\end{align}
where $\chi_m \equiv \min(\chi, \chi')$.

\end{widetext}

>>>>

\renewcommand{\section}[1]{\emph{#1}---\ignorespaces}
\renewcommand{\subsection}[1]{}

@@

    The recently discovered fast radio bursts are a new class of radio
    transient that, due to their high dispersion from intervening plasma,
    have been interpreted as
    originating from cosmological distances.  The rate of fast radio bursts
    has been estimated to be as high as 10\,000 per sky per day.  If
    cosmological, these events should trace the large-scale structure of the
    Universe in some way. Furthermore, because the dispersion measure (DM) can be
    used as a proxy for radial distance (like redshift),
    clustering can be studied in three dimensions. Due to inhomogeneities in
    the plasma's density dispersion measure is an imperfect proxy or radial
    distance and we show that this leads to additional clustering terms---the
    dispersion measure space distortions. We calculate the DM space
    power spectra for a toy model of electron and FRB clustering and show that
    the clustering signal could be detected in a survey of $10^4$ FRB
    events, as is expected to be produced by the Canadian Hydrogen Intensity
    Mapping Experiment.

@@

\section{Introduction}
The clustering of matter on large scales has been
heralded as the next great probe the Universe after the cosmic microwave
background (CMB). The large-scale structure in principle contains far more
information than the CMB, because it can be studied in three dimensions.
Traditionally the redshift of spectral lines caused by the Hubble flow has
been employed to measure position in the radial direction. Such redshift
surveys have now aggregated the positions of millions of galaxies into three
dimensional density maps exposing a rich structure of clusters, filaments,
walls and voids.

However, redshift is not the only measure of radial distance. In particular,
dispersion of electromagnetic signals due to intervening plasma
is readily measured from fast radio bursts (FRB) which are a class
of radio transient first reported in \citet{lorimer}. The dispersion measure
(DM) is a line-of-site integral of the free electron density and can thus be
used as a distance estimate.
% IF COSMOLOGICAL!!

Neither redshift or dispersion measure are perfect proxies for radial distance.
Redshifts are systematically biased by the peculiar velocities of the object
relative to the Hubble flow. This leads to additional apparent clustering 
in redshift
space which was first calculated by \citet{kaiser} using linear theory.

Similarly, distance estimates from dispersion measure will be biased by
inhomogeneities in the electron density, again leading to additional apparent
clustering of sources in dispersion measure space.  In this letter we calculate
these DM space distortions and shaw that they could 
be usefully measured in a
survey detecting as few as 10\,000 fast radio bursts.

\section{Density perturbations in dispersion measure space}
The dispersion measure of a signal observed in some angular direction $\hat{n}$ and
originating from comoving radial distance $\chi$ is
\begin{equation}
    {\rm DM}(\hat{n}, \chi) = \int_0^{\chi} \ud\chi' a(\chi)^2
        n_e(\hat{n}\chi', \chi').
\end{equation}
Here
$n_e(\vec{x}, \chi)$ is the free electron density as a function of location and
conformal time. Note that we use $\chi$ as our radial distance and time
coordinate, as opposed to redshift which is more appropriate for spectroscopic
applications. We model the cosmological 
electron density as containing a homogeneous part
and perturbations
$n_e(\vec x, \chi) = \bar{n}_e(\chi) \left[1 + \delta_e(\vec x, \chi)\right]$.
The dispersion
measure is thus
\begin{equation}
    {\rm DM}(\hat n, \chi) = \int_0^\chi \ud\chi' a(\chi')^2 \bar{n}_e(\chi')
       \left[1 + \delta_e(\hat{n}\chi',\chi')\right].
\end{equation}

@@

Combining the above equations (keeping only terms first order in the density
perturbations), we find that
\begin{equation}
\frac{\ud \chi_s}{\ud \chi} = 1 + \delta_e(\hat n \chi, \chi),
\end{equation}
and thus
\begin{equation}
\chi_s - \chi = \int_0^\chi \ud \chi' \delta_e(\hat n \chi').
\end{equation}

@@

\section{Two-point statistics}
Large-scale structure is usually studied through it two-point statistics, most
commonly the power-spectrum, $P(k)$.
Unlike in redshift-space distortions, Equation~\ref{e:delta_s} does not have a
simple form in harmonic space. The equation's third term couples harmonic modes
and thus the two-point statistics cannot be phrased as a simple power spectrum.
We will instead use $C^{ss}_\ell(\chi,\chi')$, which is
the cross-correlation angular power-spectrum of the DM space over-density
on shells at $\chi$ and $\chi'$.
\begin{align}
    &\delta_{\ell\ell'}\delta_{mm'}C^{ss}_\ell(\chi, \chi') = \nonumber \\
    & \quad
        \left\langle
        \int\ud\Omega Y_{\ell m}(\hat n)
       \delta_{fs}(\hat n \chi, \chi)
        \int\ud\Omega' Y_{\ell' m'}(\hat n')
        \delta_{fs}(\hat n' \chi', \chi')
        \right\rangle
\end{align}
The first two term in Equation~\ref{e:delta_s} are stationary. For stationary,
isotropic, tracers $x$ and $y$ we have 
$\langle \delta_x(\vec k, \chi) \delta_y(\vec k', \chi) \rangle = (2\pi)^3
\delta^3(\vec k - \vec k') P_{xy}(k, \chi)$.  If for a moment we ignore time
dependance, the angular cross power spectrum of such tracers is
\begin{align}
    %\delta_{\ell\ell'}\delta_{mm'}C^{xy}_\ell(\chi, \chi')
    %&= \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
    %    \langle \delta_x(\hat n \chi) \delta_y(\hat n' \chi') \rangle,
    %    \\
    C^{xy}_\ell(\chi,\chi')
    &= \frac{2}{\pi}
\int_0^\infty\ud k k^2 j_\ell(k\chi) j_{\ell}(k\chi')P_{xy}(k).
\end{align}
In reality the power spectrum evolves on the order of a Hubble time.
The angular cross-correlations will be very small unless $\chi$ and $\chi'$ are
within a few correlation lengths of one another, roughly a few hundred
megaparsecs.  The evolution of the power-spectrum is negligible over these time
differences which leads to a straight forward way to include the evolution of
the perturbations:
\begin{align}
C^{xy}_\ell(\chi,\chi') 
    \approx \frac{2}{\pi}
    \int_0^\infty\ud k k^2
    j_\ell(k\chi) j_{\ell}(k\chi')
    P_{xy}(k,(\chi + \chi')/2),\nonumber &\\
    |\chi - \chi'| \ll 1/aH.&
\end{align}

@@

\section{Modeling and measurement}
There are a number of effects that will complicate the measurement of the
above calculated angular power spectrum by a real survey of dispersion 
measures.  The first is
that the model that the free electron density decomposes into a homogeneous
part plus a random field of perturbations is not correct. Firstly, all signals
must pass through the Milky Way's galactic disk, halo and local environment
which will contribute to the dispersion measure. However, this is only a 
single function of $\hat n$  which can presumably be well
measured. In addition, this function could in principal be well characterized
by independent data such as observations of pulsars.  As such, the local
contribution is unlikely to be a limiting obstacle.

More concerning is the dispersion from the burst progenitor's environment
as well as the halo in which it resides.

@@

In this section we discuss some systematic effects that would affect a real
survey of dispersion measures, complicating measurement. We then estimate the
sensitivity of such a survey.

\subsection{Systematics}

@@
